Evolution of Narrow-Band Spectra in Deep-Water Constant Vorticity Flows

### 2019-04-10 Writeup/random_vorticity.tex
Quick tweak to TeX File

In non-dimensional variables, we model the evolution of a relatively small free-fluid surface over infinitely-deep water $z=\epsilon \eta(x,t)$ via the expansion 
\begin{multline*}
\eta(x,t) = \eta_{1}(\xi,\tau)e^{i\theta(x,t)} + \eta^{\ast}_{1}(\xi,\tau)e^{-i\theta(x,t)}\\
+ \epsilon\left(\eta_{0}(\xi,\tau) + \eta_{2}(\xi,\tau)e^{2i\theta(x,t)} + \eta^{\ast}_{2}(\xi,\tau)e^{-2i\theta(x,t)}\right) + \mathcal{O}\left(\epsilon^{2} \right)
\end{multline*}
where $\tau = \epsilon^{2}t$, $\xi = \epsilon(x+c_{g}t)$, and $\theta(x,t) = k_{0}x + \Omega(k_{0},\omega)t$, where the linear dispersion relationship $\Omega(k_{0},\omega)$ is given by
\[
\Omega(k_{0},\omega) = \frac{1}{2}\left(s\omega \pm \sqrt{\omega^{2} + 4|k_{0}|\left(1+\tilde{\sigma}k_{0}^{2}\right)} \right).
\]
Here, $s = \mbox{sgn}(k_{0})$, $\omega$ is the non-dimensionalized magnitude of the vorticity of the flow, and $\tilde{\sigma}$ is the surface tension.

@@

The second model we study in this paper is a higher-order generalization of the NLSE which we call the vor-Dysthe Equation (VDE) which was derived in \cite{curtis8}.  In this paper, we use it in the form 
\[
i\pd_{\tau}\eta_{1} + \alpha_{d}\pd_{\xi}^{2}\eta_{1} + \alpha_{nl}\left|\eta_{1} \right|^{2}\eta_{1} - \epsilon\frac{i(2sc_{g}\alpha_{d}+\tilde{\sigma})}{\omega-2s\Omega}\pd^{3}_{\xi}\eta_{1} + i\epsilon\mathcal{N}(\eta_{1}) = 0, 
\]
where
\[
\mathcal{N}(\eta_{1}) = \tilde{\alpha}_{1}\left|\eta_{1}\right|^{2}\pd_{\xi}\eta_{1} + \tilde{\alpha}_{2} \eta_{1}^{2}\pd_{\xi}\eta^{\ast}_{1} + i\tilde{\alpha}_{3}\eta_{1}\mathcal{H}\pd_{\xi}\left| \eta_{1}\right|^{2} + \tilde{\alpha}_{4}\pd_{\xi}\left| \eta_{1}\right|^{2}\eta_{1},
\]
and $\mathcal{H}$ is the Hilbert transform.  The terms $\tilde{\alpha}_{j}$ are straightforward to derive with the aid of a program like SAGE, but they are lengthy and not illuminative, and thus they are ommitted for the sake of brevity.

To generate simulations, we start both the NLSE and VDE from the same collection of $2L$ periodic-initial conditions
\[
\eta_{1}(\xi,0) = \epsilon_{rms}\sqrt{\frac{\delta \tilde{k}}{\sigma\sqrt{\pi}}} \sum_{k=-K+1}^{K}e^{-\tilde{k}^{2}/2\sigma^{2}}e^{i\theta_{k}}e^{i\tilde{k}\xi}, ~ \tilde{k} = \frac{\pi k}{L},
\]
where $\delta \tilde{k}=\pi/L$ and the phases $\theta_{k}$ are randomly chosen uniformly between $0$ and $2\pi$.  We can then readily show that, for $L\gg 1$, 
\[
\epsilon_{rms} \approx \overline{\left|\eta_{1}\right|^{2}}^{1/2},
\]
where $\overline{()}$ denotes the ensemble average.  In our coordinates then, the work in \cite{alber} and later confirmed in \cite{dysthe2} shows that for the NLSE the MI is suppressed for spectral widths $\sigma$ such that 
\[
\sigma \geq \epsilon_{rms}\sqrt{\frac{2\alpha_{nl}}{\alpha_{d}}}
\]
This work has helped motivate defining this parameter as the Benjamin--Feir Index (BFI), which has been shown to be a critical parameter to understanding the statstical properties of nonlinear free-surface flows; see \cite{onorato, thomas2012nonlinear}.
\section*{Results}
To understand the relationship between the vorticity $\omega$ and the initial spectral width $\sigma$ of our initial conditions, we generate ensembles of initial conditions and then run these different initial conditions up to $\tau=1$, which corresponds to $t_{f} = 1/\epsilon^{2}$ in the original fast variables.  The domain is $[-50\pi,50\pi]$.  The number of modes is $K = 256$.  The surface tension is $\tilde{\sigma} = 10^{-5}$. We then plot the averaged spectral density $S(k,t_{f};\sigma)$ where 
\[
S\left(k,t_{f};\sigma\right) = \overline{\left( \left|\frac{1}{\epsilon}\hat{\eta}_{1}\left(\frac{k-k_{0}}{\epsilon},1;\sigma \right)\right|^{2}\right)}.
\]
which corresponds to the autocorrelation of the surface profile with respect to space.  Note we have emphasized the dependence on the initial spectral width $\sigma$ for the sake of clarity.  Following the approach in \cite{dysthe2,dysthe3}, by comparing the initial spectral density to that at later times, we can see directly whether MI manifests as significant distortions of the statistics, or if through supression the distributions remain relatively close to one another.

>>>>

By expanding Equation \eqref{integro1} and taking inverse Fourier transforms, we can derive the first three terms of the corresponding Dirichlet-to-Neumann Operator (DNO) expansion, so that we find
\begin{equation}
\eta_{t} = G_{0}Q + \epsilon G_{1}Q + \epsilon^{2}G_{2}Q,
\label{dnoexp}
\end{equation}
where
\begin{align*}
G_{0}Q = & -\mathcal{H}Q\\
G_{1}Q = & \mathcal{H}\pd_{x}\left(\eta G_{0}Q\right) - \pd_{x}\left(\eta Q + \frac{\omega}{2}\eta^{2}\right)\\
G_{2}Q = & \mathcal{H}\pd_{x}\left(\eta G_{1}Q\right) + \frac{1}{2}\pd_{x}^{2}\left(\eta^{2}G_{0}Q \right) + \pd_{x}^{2}\mathcal{H}\left(\frac{1}{2}\eta^{2}Q + \frac{\omega}{3}\eta^{3}\right), 
\end{align*}
where $\mathcal{H}$ denotes the Hilbert transform so that 
\[
\mathcal{H}f = \frac{1}{2\pi}\int_{\mathbb{R}}dk~ e^{ikx}  i\mbox{sgn}(k) \hat{f}(k),
\]
where $\hat{f}(k)$ is the Fourier transform of $f(x)$ defined by
\[
\hat{f}(k) = \int_{\mathbb{R}}dx~ e^{-ikx}f(x).
\]
Using the DNO expansion in Equation \eqref{dnoexp}, we can rewrite Equation \eqref{berexp} an an evolution equation for $Q_{t}$, thereby giving us a closed system in terms of surface variables alone.  The system of equations given by Equations \eqref{dnoexp} and \eqref{berexp} we denote as the Full-Euler Equations (FEEs).

As is well known by now, in non-dimensional variables, we can model the evolution of a relatively small free-fluid surface over infinitely-deep water $z=\epsilon \eta(x,t)$ via the expansion 
\begin{multline*}
\eta(x,t) = \eta_{1}(\xi,\tau)e^{i\theta(x,t)} + \eta^{\ast}_{1}(\xi,\tau)e^{-i\theta(x,t)}\\
+ \epsilon\left(\eta_{0}(\xi,\tau) + \eta_{2}(\xi,\tau)e^{2i\theta(x,t)} + \eta^{\ast}_{2}(\xi,\tau)e^{-2i\theta(x,t)}\right) + \mathcal{O}\left(\epsilon^{2} \right)
\end{multline*}
where $\tau = \epsilon^{2}t$, $\xi = \epsilon(x+c_{g}t)$, and $\theta(x,t) = k_{0}x + \Omega(k_{0},\omega)t$, where the linear dispersion relationship $\Omega(k_{0},\omega)$ is given by
\[
\Omega(k_{0},\omega) = \frac{1}{2}\left(s\omega \pm \sqrt{\omega^{2} + 4|k_{0}|\left(1+\tilde{\sigma}k_{0}^{2}\right)} \right).
\]
Here, $s = \mbox{sgn}(k_{0})$, $\omega$ is the non-dimensionalized magnitude of the vorticity of the flow, and $\tilde{\sigma}$ is the surface tension.

@@

To generate simulations, we start both the NLSE and FEE from the same collection of $2L$ periodic-initial conditions
\[
\eta_{1}(\xi,0) = \epsilon_{rms}\sqrt{\frac{\delta \tilde{k}}{\sigma\sqrt{\pi}}} \sum_{k=-K+1}^{K}e^{-\tilde{k}^{2}/2\sigma^{2}}e^{i\theta_{k}}e^{i\tilde{k}\xi}, ~ \tilde{k} = \frac{\pi k}{L},
\]
where $\delta \tilde{k}=\pi/L$ and the phases $\theta_{k}$ are randomly chosen uniformly between $0$ and $2\pi$.  We can then readily show that, for $L\gg 1$, 
\[
\epsilon_{rms} \approx \overline{\left|\eta_{1}\right|^{2}}^{1/2},
\]
where $\overline{()}$ denotes the ensemble average.  In our coordinates then, the work in \cite{alber} and later confirmed in \cite{dysthe2} shows that for the NLSE the MI is suppressed for spectral widths $\sigma$ such that 
\[
\sigma \geq \epsilon_{rms}\sqrt{\frac{2\alpha_{nl}}{\alpha_{d}}}
\]
This work has helped motivate defining this parameter as the Benjamin--Feir Index (BFI), which has been shown to be a critical parameter to understanding the statstical properties of nonlinear free-surface flows; see \cite{onorato, thomas2012nonlinear}.
\section*{Results}
To understand the relationship between the vorticity $\omega$ and the initial spectral width $\sigma$ of our initial conditions, we generate ensembles of initial conditions and then run these different initial conditions up to $\tau=1$, which corresponds to $t_{f} = 1/\epsilon^{2}$ in the original fast variables.  The domain is $[-50\pi,50\pi]$.  The number of modes is $K = 256$.  The surface tension is $\tilde{\sigma} = 10^{-5}$. We then plot the averaged spectral density $S(k,t_{f};\sigma)$ where 
\[
S\left(k,t_{f};\sigma\right) = \overline{\left( \left|\frac{1}{\epsilon}\hat{\eta}_{1}\left(\frac{k-k_{0}}{\epsilon},1;\sigma \right)\right|^{2}\right)}.
\]
which corresponds to the autocorrelation of the surface profile with respect to space.  Note we have emphasized the dependence on the initial spectral width $\sigma$ for the sake of clarity.  Following the approach in \cite{dysthe2,dysthe3}, by comparing the initial spectral density to that at later times, we can see directly whether MI manifests as significant distortions of the statistics, or if through supression the distributions remain relatively close to one another.
